Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-08-04 PAPER.tex
-Made Aaron changes for first half of paper

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the birth of the first stars and galaxies led to the ionization of neutral hydrogen. This important epoch of our cosmic dawn harbors a wealth of information regarding the environment during this transformative time, including insight into the nature of the first luminous sources and implications about cosmological parameters. There are many experiments investigating the EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to develop techniques that help maximize sensitivity and validate results. Additionally, it is imperative to understand the trade-offs between different methods and their effects on common power spectrum themes. In this paper, we focus on three major themes - signal loss, power spectrum error bar estimation, and bias in measurements. We describe techniques that affect these themes using both a toy model and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER).
\end{abstract}

@@

By about one billion years after the Big Bang, the very first stars and galaxies are thought to have ionized all the neutral hydrogen that dominated the baryonic matter content in the early Universe. This transition period, during which the first luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas into a plasma, is known as the epoch of reionization (EoR). The EoR is a relatively unexplored era in our cosmic dawn. Its history encodes important information regarding the nature of the first galaxies and the processes of structure formation. Direct measurements of the EoR would unlock powerful information about the intergalactic medium, revealing connections between the smooth matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured web of galaxies we observe today.

One promising technique to probe the EoR is to target the 21 cm wavelength emission that is emitted by neutral hydrogen via its spin-flip transition. This technique is powerful because it can be observed as a function of redshift --- that is, the wavelength of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before stretching out as it traveled through expanding space. The 21 cm line therefore offers a window into the evolution of ionization, temperature, and density fluctuations on cosmic scales.

Although a detection of the EoR remains elusive, there are several radio telescope experiments that have succeeded in using the 21 cm signal from hydrogen to place constraints on the brightness of the EoR. Examples of experiments investigating the mean brightness temperature of the EoR relative to the CMB are the Experiment to Detect the Global EoR Signature (EDGES, \citealt{bowman2010}), the Long Wavelength Array (LWA, \citealt{ellingson_et_al2009}),  the Large Aperture Experiment to Detect the Dark Ages (LEDA, \citealt{greenhill_bernardi2012}), the Dark Ages Radio Explorer (DARE, \citealt{burns2012}), the Sonda Cosmol\'ogica de las Islas para la Detecci\'on de Hidr\'ogeno NeutroSciHi (SCI-HI, \citealt{voytek2014}), the Broadband Instrument for Global HydrOgen ReioNisation Signal (BIGHORNS, \citealt{sokolowski2015}), and the Shaped Antenna measurement of the background RAdio Spectrum (SARAS, \citealt{patra2015}). Radio interferometers which seek to measure statistical power spectra include the Giant Metre-wave Radio Telescope (GMRT, \citealt{paciga_et_al2013}), the LOw Frequency ARray (LOFAR,\citealt{van_haarlem_et_al2013}), the Murchison Widefield Array (MWA, \citealt{tingay_et_al2013}), the 21 Centimeter Array (21CMA, \citealt{peterson_et_al2004}, \citealt{wu2009}), and PAPER (\citealt{parsons_et_al2010}). The Hydrogen Epoch of Reionization Array (HERA), which is currently being built, is a next-generation instrument that aims to combine lessons learned from previous experiments and is forecasted to be able to make a successful $23\sigma$ detection with an eventual $350$ elements, using current analysis techniques (\citealt{deboer_et_al2017}, \citealt{pober_et_al2014}).

The major challenge that faces all 21 cm experiments is isolating a small signal that is buried underneath foregrounds and instrumental systematics that are 4-5 orders of magnitude brighter. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact with each other to affect power spectrum results. In this paper, we discuss three themes essential to a $21$ cm power spectrum analysis and how data analysis choices affect each. We approach these themes from a broad perspective, and then perform a detailed case study using data from the 64-element configuration of PAPER.

This paper is organized as follows. In section 2 we introduce the three themes of our focus, using a toy model to develop intuition into each one. Section 3 presents an overview of the PAPER-64 array and observations, highlighting key changes from \citet{ali_et_al2015}. Sections 4, 5, and 6 detail how the new PAPER-64 analysis quantifies signal loss, estimates error bars, and eliminates bias, respectively. We conclude in Section 7.

@@

There are many choices for a data analyst, such as how to combine time-ordered measurements, how to estimate its variance, how to weight data in a way that suppresses contaminated modes while not destroying an EoR signal, and how to identify the source of a detection. Many common techniques, such as averaging data, weighting, bootstrapping, and jack-knife testing, each address these issues yet harbor additional trade-offs. For example, an aggressive filtering method may succeed in eliminating interfering systematics but comes at the cost of losing some EoR signal. A chosen weighting scheme may maximize sensitivity but fail to suppress foregrounds.

Measuring the statistical $21$ cm power spectrum requires robust methods for determining accurate confidence intervals and rigorous techniques to identify and suppress systematics. Additionally, we desire to maximize our sensitivity with filtering and weighting techniques. In this paper, we focus on four techniques and their effects on three overarching themes of $21$ cm power spectrum estimation. We will give brief definitions now, and build intuition for each theme in the sections to follow.

\begin{center}
Power Spectrum Themes
\end{center}
\begin{itemize}
\item[] The accuracy and interpretation of a $21$ cm power spectrum can be encompassed in the following three themes.
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): As explained in the next section, certain analysis techniques can lead to the loss of the EoR signal. If not corrected for, it could lead to a false non-detection. Computing signal loss has subtle challenges but is crucial to the accuracy of any result.
\item \textbf{Error Bar Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on the $21$ cm power spectrum result determine the difference between a detection and a null result, which have two very different implications. Errors can be estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of bias in a visibility measurement that can leak its way into a power spectrum, such as bias from noise and foregrounds. Proving a bias is EoR may be the most difficult challenge for $21$ cm analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper we will highlight some sources of bias and discuss ways to mitigate their effects.
\end{itemize}

\begin{center}
Power Spectrum Techniques
\end{center}
\begin{itemize}
\item[] The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in data. However, they also have limitations and circumstances in which there are trade-offs.
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering can be considered an averaging scheme for time-ordered data (\citealt{parsons_et_al2016}), and we explain the trade-offs of filtering in more detail in Section \ref{sec:SiglossOverview}. Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence of foregrounds and systematics. 
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain spectral features and minimize others. One particular flavor of weighting is inverse covariance weighting, which weights a dataset by minimizing the covariance between frequency channels. This weighting has the effect of down-weighting correlated information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, one challenge of weighting is in accurately determining a covariance matrix that best describes the data.
\item \textbf{Bootstrapping:} Because applying theoretical errors to a $21$ cm measurement are insufficient \cc{why?}, bootstrapping is a useful method for estimating errors of a dataset from itself. By randomly drawing many samples of the data, we get a sense of its inherent variance, though there are subtleties to consider such as bootstrapping sample sizes.
\item \textbf{Jack-Knife testing:} A resampling technique useful for estimating bias, jack-knives can be taken along different dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of systematics (\citealt{keating_et_al2016}).
\end{itemize}

@@

Signal loss refers to the attenuation of the target cosmological signal in a dataset. It can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during spectral calibration, applying a delay-domain filter, or by weighting data by itself. Here we focus on signal loss associated with applying a weighting matrix to data. Driven by the need to mitigate foreground bias, we would like to use a weighting method that succeeds in down-weighting foregrounds. However, as we'll investigate, this carries the risk of large amounts of signal loss.

Before we demonstrate how different weighting matrices affect signal loss, we first recap how weights are related to the power spectrum estimation technique of optimal quadratic estimators (OQE, \citealt{liu_et_al2014b}). A summary of the method is as follows.

We begin with our data vector, $\textbf{x}$, which contains our measured visibilities in mK that have the shape ($times$, $frequencies$). We form the un-normalized power spectrum estimate $\hat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\hat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{w}\textbf{Q}_{\alpha}\textbf{w}\textbf{x}
\end{equation}

where $\textbf{w}$ \cc{change to capital letter that isn't taken yet!} is a weighting matrix (for example, inverse covariance weighting would set $\textbf{w} = \textbf{C}^{-1}$). The matrix $\textbf{Q}$ is an operator that takes our frequency-domain visibilities and Fourier-transforms them into power spectrum space. The index $\alpha$ denotes a waveband in $k_{\parallel}$, where $k_{\parallel}$ is the Fourier-dual to frequency.

@@

\begin{equation}
\hat{\textbf{p}} = \textbf{M}\hat{\textbf{q}}
\end{equation}

where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. We compute $\textbf{M}$ using the Fisher matrix $\textbf{F}$, defined as:

\begin{eqnarray}
\textbf{F}_{\alpha\beta} &=& \frac{1}{2}tr[\textbf{w}\textbf{Q}_{\alpha}\textbf{w}\textbf{Q}_{\beta}]. 
\end{eqnarray}

We have a choice for $\textbf{M}$, and for simplicity in this section we set $\textbf{M} \propto \textbf{I}$, the identity matrix. This is not the case for the analysis of PAPER-64 data, which we describe in Section \cc{not sure which section to point to yet} as part of the discussion of bias.

In the next sections, we will experiment with different weighting matrices $\textbf{w}$ and their impact on the resulting power spectra estimate $\hat{\textbf{p}}$.

\subsubsection{Toy Model: Inverse Covariance Weighting}

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we use a toy model. Suppose we have a very simple dataset that contains 2-dimensional data (representing $100$ time integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which could be used for a power spectrum analysis. We create a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid that varies in time and frequency, as well as a mock EoR signal, $\textbf{x}_{EoR}$, as complex, random Gaussian noise. The sum of the two forms our combined data vector, $\textbf{x}$, to which we will be applying different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}. Our ultimate goal in experimenting with weighting is to suppress the foregrounds perfectly without destroying EoR.

One choice for a weight matrix $\textbf{w}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it is an aggressive way to down-weight foregrounds. The covariance matrix $\textbf{C}$, which describes covariances between frequency channels, can be estimated in a variety of ways. Should we have perfect foreground, instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our real data. However, if our foreground model is flawed for example, our estimate of $\textbf{C}$ would not be successful at down-weighting the real foregrounds in the data. One attractive way to estimate $\textbf{C}$ then is to empirically derive it from the data vector $\textbf{x}$ itself:

\begin{equation}
\textbf{C} \equiv \langle\textbf{xx}^{\dagger}\rangle,
\end{equation}

assuming $\langle\textbf{x}\rangle = 0$. The inverse covariance matrix is therefore $\textbf{w} = \textbf{C}^{-1}$.

If $\textbf{C}$ is computed from the data itself and its inverse is then applied to the data, it carries the risk of over-fitting information in the data and introducing a multiplicative bias to estimates of the signal. If instead we choose $\textbf{w} = \textbf{I}$, the identity matrix, there will be no signal loss (although it may suffer from foreground or instrumental bias). We will refer to this choice as `unweighted', since using the identity does not alter (doesn't upweight or downweight) the data at all.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss1.png}
	\caption{Our toy model dataset contains a sinusoid foreground that varies in time and frequency and random Gaussian noise as a mock EoR signal. Real parts are shown here. \cc{colorbar?}}
	\label{fig:toy_sigloss1}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 1cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={1cm 0.2cm 0cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for our toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data (left), contrasted with projecting out the first eigenmode only (right). There is signal loss for the former but not the latter.}
	\label{fig:toy_sigloss3}
\end{figure*}

As shown in Figure \ref{fig:toy_sigloss3} (left plot), our inverse covariance weighted result successfully suppresses foregrounds. This illustrates why inverse covariance weighting is attractive for $21$ cm data. It is also evident that our result fails to recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low. This is evidence of signal loss. In order to understand how this occurs, we need to closely study our covariance matrix, $\textbf{C}$.

It turns out that because we estimated $\textbf{C}$ from our data, its eigenspectrum offers insight into understanding how inverse covariance weighting affects our result. An eigenspectrum ranks the eigenvalues of $\textbf{C}$ from highest to lowest and can be thought of as a spectrum of weights that are given to each frequency mode in the data. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of foregrounds only (blue), EoR only (red), and FG + EoR (green). The eigenspectrum of foregrounds only peaks at the zeroth eigenmode, a consequence of the presence of only one sinusoid. All eigenspectra steepen slightly for the final few eigenmodes, a behavior that results in some modes being upweighted more strongly than others, thereby overfitting noise and causing signal loss in the final power spectrum estimate (Figure \ref{fig:toy_sigloss3}, left). If only the first eigenmode is used, all others get equal weights and hence, there is no signal loss (Figure \ref{fig:toy_sigloss3}, right)}
	\label{fig:toy_sigloss2}
\end{figure}

For our toy model, the eigenspectra of $\textbf{C}_{FG}$, $\textbf{C}_{EoR}$, and $\textbf{C}$ are shown in Figure \ref{fig:toy_sigloss2}. The blue curve peaks only at the zeroth eigenmode, revealing that the foregrounds can be described by a single eigenmode. This is a direct consequence of how it was constructed (only one sinusoid). In fact, the blue peak in Figure \ref{fig:toy_sigloss3} is also a direct result of our foreground's construction. The eigenspectrum of EoR, however, is fairly flat, signifying that its covariance matrix is similar to an identity matrix. This makes sense because the EoR signal consists of random noise, which should only have strong frequency-frequency correlations when two frequencies are the same.

The eigenmodes with the highest eigenvalues will be down-weighted the most (i.e. when the spectrum is inverted, they will be given the smallest weights). In contrast, the modes with the smallest eigenvalues will be up-weighted the most. In other words, if an eigenspectrum is not perfectly flat, all modes will be given different weights and it is possible to up-weight a few modes much more than others. The subtle slope of the eigenspectrum in this toy example gives rise to a small amount of signal loss because it is possible to overfit a few modes (the highest weighted ones) ever so slightly. We will see in the next section how this effect can be exaggerated as the `steepness' of the spectrum changes.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Namely, we can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. The result is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case, we perfectly recover EoR, demonstrating that if we can perfectly model our foregrounds, we can down-weight them without signal loss. However, this is not the case in reality since we aren't able to distinguish between foreground and EoR modes.

>>>>

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the birth of the first stars and galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to develop techniques that help maximize sensitivity and validate results. Additionally, it is imperative to understand the trade-offs between different methods and their effects on common power spectrum themes. In this paper, we focus on three major themes --- signal loss, power spectrum error bar estimation, and bias in measurements. We describe techniques that affect these themes using both a toy model and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER). \cc{motivate a bit more once the paper is complete}
\end{abstract}

@@

By about one billion years after the Big Bang ($z \sim 6$), the very first stars and galaxies are thought to have ionized all the neutral hydrogen that dominated the baryonic matter content in the Universe. This transition period, during which the first luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas into a plasma, is known as the epoch of reionization (EoR). The EoR is a relatively unexplored era in our cosmic dawn. Its history encodes important information regarding the nature of the first galaxies and the processes of structure formation. Direct measurements of the EoR would unlock powerful information about the intergalactic medium, revealing connections between the smooth matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured web of galaxies we observe today.

One promising technique to probe the EoR is to target the 21 cm wavelength emission that is emitted by neutral hydrogen via its spin-flip transition. This technique is powerful because it can be observed as a function of redshift --- that is, the wavelength of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before stretching out as it traveled through expanding space. The 21 cm line offers a window into the evolution of ionization, temperature, and density fluctuations on cosmic scales.

Although a detection of the EoR remains elusive, there are several radio telescope experiments that have succeeded in using the 21 cm signal from hydrogen to place constraints on the brightness of the EoR. Examples of experiments investigating the mean brightness temperature of the EoR relative to the CMB are the Experiment to Detect the Global EoR Signature (EDGES; \citealt{bowman2010}), the Long Wavelength Array (LWA; \citealt{ellingson_et_al2009}),  the Large Aperture Experiment to Detect the Dark Ages (LEDA; \citealt{greenhill_bernardi2012}), the Dark Ages Radio Explorer (DARE; \citealt{burns2012}), the Sonda Cosmol\'ogica de las Islas para la Detecci\'on de Hidr\'ogeno NeutroSciHi (SCI-HI; \citealt{voytek2014}), the Broadband Instrument for Global HydrOgen ReioNisation Signal (BIGHORNS; \citealt{sokolowski2015}), and the Shaped Antenna measurement of the background RAdio Spectrum (SARAS; \citealt{patra2015}). Radio interferometers which seek to measure statistical power spectra include the Giant Metre-wave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), the LOw Frequency ARray (LOFAR;\citealt{van_haarlem_et_al2013}), the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the 21 Centimeter Array (21CMA; \citealt{peterson_et_al2004}; \citealt{wu2009}), and PAPER (\citealt{parsons_et_al2010}). The Hydrogen Epoch of Reionization Array (HERA), which is currently being built, is a next-generation instrument that aims to combine lessons learned from previous experiments and is forecasted to be able to make a successful high-significance power spectrum detection with an eventual $350$ elements using current analysis techniques (\citealt{deboer_et_al2017}; \citealt{pober_et_al2014}).

The major challenge that faces all 21 cm experiments is isolating a small signal that is buried underneath foregrounds and instrumental systematics that are 4-5 orders of magnitude brighter. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure the accuracy and reliability of a potential EoR detection. In this paper, we discuss three themes that are essential to investigate for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, traps, and connections to the themes . We first approach the themes from a broad perspective, and then perform a detailed case study using data from the 64-element configuration of PAPER.

This paper is organized as follows. In section 2 we introduce the three themes of our focus, using a toy model to develop intuition into each one. Section 3 presents a case study into each theme using data from the PAPER-64 array, highlighting key changes from \citet{ali_et_al2015}. We conclude in Section 7.

@@

There are many choices for a data analyst. How can time-ordered measurements be combined? How can the variance of the data be estimated? In what way(s) can the data be weighted to suppress contaminated modes while not destroying an EoR signal? How can the source of a detection be properly identified? Many common techniques, such as averaging data, weighting, bootstrapping, and jackknife testing, each address these issues yet harbor additional trade-offs. For example, an aggressive filtering method may succeed in eliminating interfering systematics but comes at the cost of losing some EoR signal. A chosen weighting scheme may maximize sensitivity but fail to suppress foregrounds.

Despite there being many data analysis choices, measuring the statistical $21$ cm power spectrum ultimately requires robust methods for determining accurate confidence intervals and rigorous techniques to identify and suppress systematics.  In this paper, we focus on three $21$ cm power spectrum themes that encapsulate this goal and discuss four techniques that interplay with each other and the themes. We will give brief definitions now, and build intuition for each theme in the sections to follow.

\begin{center}
Power Spectrum Themes
\end{center}

A deep understanding of the following three themes is essential for the accuracy and interpretation of a $21$ cm power spectrum result. Stemmed from a re-analysis of PAPER-64 data, we believe these themes serve as an important check-list for a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): As explained in the next section, certain analysis techniques can lead to the loss of the EoR signal. If not corrected for, it could lead to a false non-detection. Computing signal loss has subtle challenges but is crucial to the accuracy of any result.
\item \textbf{Error Bar Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on the $21$ cm power spectrum result determine the difference between a detection and a null result, which have two very different implications. Errors can be estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of bias in a visibility measurement that can leak its way into a power spectrum, such as bias from noise and foregrounds. Proving a bias is EoR may be the most difficult challenge for $21$ cm analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper we will highlight some sources of bias and discuss ways to mitigate their effects.
\end{itemize}

\begin{center}
Power Spectrum Techniques
\end{center}

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in data. We focus on these techniques because of their strong appeal for a power spectrum analysis. However, they also have limitations, and we will discuss circumstances in which there are trade-offs.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering can be considered an averaging scheme for time-ordered data (\citealt{parsons_et_al2016}), and we explain the trade-offs of filtering in more detail in Section \ref{sec:SiglossOverview}. Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence of foregrounds and systematics. 
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain spectral features and minimize others. One particular flavor of weighting is inverse covariance weighting, which weights a dataset by minimizing the covariance between frequency channels. This weighting has the effect of down-weighting correlated information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, one challenge of weighting is in accurately determining a covariance matrix that best describes the data.
\item \textbf{Bootstrapping:} Because applying theoretical errors to a $21$ cm measurement are insufficient \cc{why?}, bootstrapping is a useful method for estimating errors of a dataset from itself. By randomly drawing many samples of the data, we get a sense of its inherent variance, though there are subtleties to consider such as bootstrapping sample sizes.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of systematics (\citealt{keating_et_al2016}).
\end{itemize}

@@

In this section we illustrate how signal loss arises in analysis. Signal loss refers to attenuation of the target cosmological signal in a power spectrum estimate. It can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during spectral calibration, applying a delay-domain filter, or by weighting data by itself. Here we focus on signal loss associated with applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was previously underestimated in the PAPER-64 analysis.

Driven by the need to mitigate foreground bias, we would like to use a weighting method that succeeds in down-weighting foregrounds. We apply this weighting to data and propagate it into a final estimator using the power spectrum estimation technique of optimal quadratic estimators (OQE; \citealt{liu_tegmark2011}; \citealt{liu_et_al2014b}). Before showing how signal loss can arise when using different weighting matrices, we first summarize OQE as follows.

We begin with our data vector, $\textbf{x}$, which contains our measured visibilities in Jy that have the shape ($times$, $frequencies$). We form the un-normalized power spectrum estimate $\hat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\hat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{S}\textbf{Q}_{\alpha}\textbf{S}\textbf{x},
\end{equation}

\noindent where $\textbf{S}$ is a weighting matrix (we use $\textbf{S}$ since $\textbf{W}$ is reserved for window functions). As an example, inverse covariance weighting would set $\textbf{S} = \textbf{C}^{-1}$ and an unweighted case would use $\textbf{S} = \textbf{I}$, the identity matrix. The matrix $\textbf{Q}$ is an operator that takes our frequency-domain visibilities and Fourier-transforms them into power spectrum space. The index $\alpha$ denotes a waveband in $k_{\parallel}$, where $k_{\parallel}$ is the Fourier-dual to frequency.

@@

\begin{equation}
\hat{\textbf{p}} = \textbf{M}\hat{\textbf{q}},
\end{equation}

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. We compute $\textbf{M}$ using the Fisher matrix $\textbf{F}$, defined as:

\begin{eqnarray}
\textbf{F}_{\alpha\beta} &=& \frac{1}{2}tr[\textbf{S}\textbf{Q}_{\alpha}\textbf{S}\textbf{Q}_{\beta}]. 
\end{eqnarray}

We have a choice for $\textbf{M}$. For simplicity in this section we set $\textbf{M} \propto \textbf{I}$, although we explore other cases for the analysis of PAPER-64 data as explained in Section \ref{sec:PS}

In the next sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{S}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds perfectly without destroying an EoR signal.

\subsubsection{Toy Model: Inverse Covariance Weighting}
\label{sec:toymodel}

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which might be used for a power spectrum analysis. We create a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid that varies in time and frequency (\cc{explain why a sinusoid for FG}), as well as a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. Our combined data vector is $\textbf{x} = \textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss1.png}
	\caption{Our toy model dataset contains a sinusoid foreground that varies in time and frequency and a random Gaussian signal as a mock EoR signal. Real parts are shown here. \cc{colorbar?}}
	\label{fig:toy_sigloss1}
\end{figure}

One choice for a weight matrix $\textbf{S}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it is an aggressive way to down-weight foregrounds. \cc{math!} The covariance matrix $\textbf{C}$, which in our case describes covariances between frequency channels, can be estimated in a variety of ways. Given perfect foreground, instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data. \cc{math!} However, if our foreground model is flawed for example, our estimate of $\textbf{C}$ would not be successful at down-weighting the foregrounds in the data.

One attractive way to estimate $\textbf{C}$ is to empirically derive it from the data vector $\textbf{x}$ itself:

\begin{equation}
\hat{\textbf{C}} \equiv \langle\textbf{xx}^{\dagger}\rangle,
\end{equation}

\noindent assuming $\langle\textbf{x}\rangle = 0$, where $\langle \rangle$ denotes a time-average \cc{distinguish between ideal ensemble average and finite estimate???}. The inverse covariance matrix is therefore $\textbf{S} = \textbf{C}^{-1}$. \cc{should I put hat's on all C's that are estimated from the data? Need to distinguish it from ideal C.}

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 1cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={1cm 0.2cm 0cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymode} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data (left), contrasted with projecting out the first eigenmode only (right). There is signal loss for the former but not the latter. \cc{elaborate}}
	\label{fig:toy_sigloss3}
\end{figure*}

As shown, our inverse covariance-weighted result successfully suppresses foregrounds \cc{explain why}. This illustrates why inverse covariance weighting is attractive for $21$ cm data. It is also evident that our result fails to recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low. This is evidence of signal loss. In order to understand how this occurs, we need to closely study our covariance matrix, $\textbf{C}$, which is shown in Figure \ref{fig:toy_sigloss12}.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 2.3cm 0.3cm 2.3cm},clip,width=\columnwidth]{plots/toy_sigloss12.png}
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss13.png}
	\caption{The covariance matrices (top row) and inverse covariance-weighted data (bottom row) for FG only, EoR only, and FG + EoR. Real parts are shown here. \cc{colorbar?}}
	\label{fig:toy_sigloss12}
\end{figure}

If $\textbf{C}$ is computed from the data itself and its inverse is then applied to the data, it carries the risk of over-fitting information in the data and introducing a multiplicative bias to estimates of the signal \cc{elaborate with math!}. As a contradictory example, if instead we choose $\textbf{S} = \textbf{I}$, there will be no signal loss (although it may suffer from foreground or instrumental bias).

It turns out that because we estimated $\textbf{C}$ from our data, its eigenspectrum offers insight into understanding how inverse covariance weighting affects our result. An eigenspectrum ranks the eigenvalues of $\textbf{C}$ from highest to lowest and can be thought of as a spectrum of weights that are given to each frequency mode in the data. Each eigenvalue encodes the amount of variance in the dataset for a particular mode. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. \cc{use more rigorous math to supplement this paragraph}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\textbf{C}_{FG}$ (blue), $\textbf{C}_{EoR}$ (red), and $\textbf{C}_{FG+EoR}$ (green). The eigenspectrum of $\textbf{C}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid. All eigenspectra steepen slightly for the final few eigenmodes, a behavior that results in some modes being upweighted more strongly than others, thereby overfitting noise and causing signal loss in the final power spectrum estimate (Figure \ref{fig:toy_sigloss3}, left). If only the first eigenmode is used, all others get equal weights and hence, there is no signal loss (Figure \ref{fig:toy_sigloss3}, right). \cc{also illustrate first few and last few eigenvalues -- how to best show this?}}
	\label{fig:toy_sigloss2}
\end{figure}

For our toy model, the eigenspectra of $\textbf{C}_{FG}$, $\textbf{C}_{EoR}$, and $\textbf{C}$ are shown in Figure \ref{fig:toy_sigloss2}. The blue curve peaks only at the zeroth eigenmode, revealing that the foregrounds can be described by a single eigenmode. This is a direct consequence of how it was constructed (only one sinusoid). The eigenspectrum of EoR, however, is fairly flat, signifying that its covariance matrix is similar to an identity matrix. This makes sense because EoR consists of a random signal, which should only have strong frequency-frequency correlations when two frequencies are the same.

The eigenmodes with the highest eigenvalues will be down-weighted the most; when the spectrum is inverted, they will be given the smallest weights. In contrast, the modes with the smallest eigenvalues will be up-weighted the most. In other words, if an eigenspectrum is not perfectly flat, all modes will be given different weights and it is possible to up-weight a few modes much more than others. The subtle slope of the eigenspectrum in this toy example is the result of describing shapes in the data that aren't physically there \cc{reword sentence... the shape is there but it's not valid?}. This gives rise to a small amount of signal loss because it is possible to overfit a few modes (the highest weighted ones) ever so slightly. We will see in the next section how this effect can be exaggerated as the `steepness' of the spectrum changes.
